---
title: "Singapore Trade Balance - How Covid-19 has affected the landscape"
subtitle: "Take Home Exercise 4"
author: "Wang Kunrui"
institute: "School of Computing and Information Systems,<br/>Singapore Management University"
date: "2023-2-26"
output:
  xaringan::moon_reader:
    css: [xaringan-themer.css, "css/text.css"]
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      beforeInit: "js/macros.js"
---

# Overview

The goal of this assignment is to uncover how changes in the global economical and political dynamics, in lieu of COVID-19, has impacted Singapore's bilateral trade. For the purpose of this study, we used the data "Merchandise Trade by Region/Market" from Department of Statistics, available [here](https://www.singstat.gov.sg/find-data/search-by-theme/trade-and-investment/merchandise-trade/latest-data)

# Data Wrangling

The following steps were required for Data Wrangling;

1.  Narrow the scope of study to Covid-related period: 2020 to 2022, plus 2019 for comparison purpose

2.  Align the trade value to be based in "Million Dollars" for simple comparison

3.  Create a separate document with merged values (by Trade Partner) in Import and Export

```{r, include=FALSE}
#| warning: false

#create base dataframe
library(pacman)
pacman::p_load(ggstatsplot, ggplot2, gapminder, ggside, knitr, tidyverse, broom, ggdist, gganimate, plotly, crosstalk, DT, hrbrthemes, viridis,realxl,gganimate,gifski,png,stringr,lubridate,lattice, shiny, rmarkdown)
Import <- readxl::read_xlsx("raw.xlsx",sheet="T1")
Export <- readxl::read_xlsx("raw.xlsx",sheet="T2")
names(Import)<- Import[9,]
names(Export)<- Export[9,]
Import <- Import [10:128,]
Export <- Export [10:100,]
```

```{r, include=FALSE}
#creating code chunk for first chart
Chart1_Import <- Import %>% 
  mutate (across (-c(`Data Series`), as.numeric)) %>%
  rowwise%>% mutate (FY19 = sum(c_across(contains("2019")))) %>%
  rowwise%>% mutate (FY20 = sum(c_across(contains("2020")))) %>%
  rowwise%>% mutate (FY21 = sum(c_across(contains("2021")))) %>%
  rowwise%>% mutate (FY22 = sum(c_across(contains("2022")))) %>%
  select ("Data Series"|starts_with("FY"))

Chart1_Export <- Export %>% 
  mutate (across (-c(`Data Series`), as.numeric)) %>%
  rowwise%>% mutate (FY19 = sum(c_across(contains("2019")))) %>%
  rowwise%>% mutate (FY20 = sum(c_across(contains("2020")))) %>%
  rowwise%>% mutate (FY21 = sum(c_across(contains("2021")))) %>%
  rowwise%>% mutate (FY22 = sum(c_across(contains("2022")))) %>%
  select ("Data Series"|contains("FY"))

Chart1 <- merge (Chart1_Export,Chart1_Import, by="Data Series")
Chart1 <- Chart1 %>%
  rename_with(~gsub(".x",".Export",.)) %>%
  rename_with(~gsub(".y",".Import",.)) %>%
  rowwise %>% mutate (FY19net = FY19.Export-FY19.Import) %>%
  rowwise %>% mutate (FY20net = FY20.Export-FY20.Import) %>%
  rowwise %>% mutate (FY21net = FY21.Export-FY21.Import) %>%
  rowwise %>% mutate (FY22net = FY22.Export-FY22.Import) %>%
  mutate(across(-`Data Series`, ~ifelse(str_detect(`Data Series`,"Thousand"),./1000,.))) %>%
  mutate(`Data Series`= gsub("(Thousand Dollars)","",`Data Series`)) %>%
  mutate(`Data Series`= gsub("(Million Dollars)","",`Data Series`))

Chart1 <- Chart1 %>%
  filter(!`Data Series` %in% c("Asia ()","America ()","European Union ()","Africa ()","Oceania ()","Other Countries In Oceania ()","Europe ()"))


```

# Animated overview of top trading partners

For the scope of this study, we want to observe any changes in trade dynamic between Singapore and its Top 10 Export Partners pre-covid (i.e. in 2019) and Top 10 Import Partners.

Singapore's Top Export Partners in 2019 are:

```{r, echo=FALSE}
topExport <- head(Chart1 %>%
  select ("Data Series"|"FY19.Export") %>%
  arrange(desc(FY19.Export)),
  10)
topImport <- head(Chart1 %>%
  select ("Data Series"|"FY19.Import") %>%
  arrange(desc(FY19.Import)),
  10)
top<-c(topExport$`Data Series`,topImport$`Data Series`)
top<-unique(top)
print ((topExport$`Data Series`))
```

Singapore's Top Import Partners in 2019 are:

```{r, echo=FALSE}
print ((topImport$`Data Series`))
```

Quick Fact:

"A nation's net exports are the value of its total exports minus the value of its total imports." - Investopedia

Chart 1 below showcase the Net Export for shortlisted partners above.

The following steps were required for further Data Wrangling;

1.  Created a calculated column for Net Export

2.  Retain only the Top 10 Import Partners and Top Export Partners

### Key observations from this table:

-   Singapore's top net trading partner is Hong Kong, and our bottom net trading partner in Taiwan. This remained consistent across the years pre- and post- pandemic (i.e. 2019 to 2022).

-   We observe changes in trade dynamics with the following partners during the pandemic:

    -   Mainland China

    -   Republic of Korea

    -   United States

```{r, include=FALSE}

#plot for chart 1
Plot1 <- Chart1 %>%
  select ("Data Series"|contains("net")) %>%
  arrange(desc(FY19net))
top<-head(Plot1,10)
bottom<-tail(Plot1,10)
Plot1 <- rbind(top,bottom)
```

```{r echo=FALSE}

Plot1a<-Plot1%>%
  pivot_longer(!`Data Series`,names_to = "year", values_to = "ttl")
  
Plot1a<- Plot1a %>%
  mutate(`Data Series`= gsub("()","",`Data Series`)) %>%
  mutate(year= gsub("net","",year)) %>%
  mutate(year= gsub("FY","20",year))

p<-ggplot(Plot1a, aes(fct_reorder(`Data Series`,-ttl),y=ttl,color=ttl>0))+
  geom_col()+
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))+
  labs(subtitle="{closest_state} Net Export with Top Trading Partners")+
  theme(legend.position = "none")


Countries<-c(top$`Data Series`, bottom$`Data Series`)

anim <- p+transition_states(year,transition_length = 2,state_length = 1)+
  xlab("Trading Partner") + 
  ylab("Total Value in SGD (million)")

anim
```

# Comprehensive view of volume between 2020 - 2022, by trading partner

The trellis plot below provides details of trade volume between Singapore and Individual top and bottom Net Export Trade Partners, This chart is powerful as it enables us to view trends - across time - in a singular view. The mouse-over (hover) function also provides users with details on exact value transacted.

The following steps were required for further Data Wrangling;

1.  Pivot_longer on earlier dataframes to create columns for Dates and Trade Type (Import or Export)

2.  Data wrangling on date format using stringr package

3.  Ensure that variable to identify country (Data Series in this project) is a factor which can be identified for trellis plot

### Key observations from this table:

-   Our top trading partners are:

    -   Hong Kong

    -   Taiwan

    -   United States

    -   Indonesia

-   Monthly trade (either Import or Export) exceeding the value of SGD 2B is observed to be higher than other trading partners.

-   Exports to several countries (Indonesia, Taiwan) increased gradually during the pandemic. For example, exports to Indonesia grew steadily from 1.7B SGD in May-2020 to 5.3B SGD in July 2022.

-   Softening of trade between Singapore and its top partners (aside from Hong Kong) was observed, beginning Q3 last year.

```{r, include=FALSE}
Chart2 <- merge (Export,Import, by="Data Series")
Chart2 <- Chart2 %>%
  filter (.[[2]]!=0) %>%
  mutate (across (-c(`Data Series`), as.numeric)) %>%
  mutate(across(-`Data Series`, ~ifelse(str_detect(`Data Series`,"Thousand"),./1000,.))) %>%
  mutate(`Data Series`= gsub("(Thousand Dollars)","",`Data Series`)) %>%
  mutate(`Data Series`= gsub("(Million Dollars)","",`Data Series`)) %>%
  filter(!`Data Series` %in% c("Asia ()","America ()","European Union ()","Africa ()","Oceania ()","Other Countries In Oceania ()","Europe ()")) %>%
  select ("Data Series"|contains("2020")|contains("2021")|contains("2022")) %>%
  pivot_longer(!`Data Series`,names_to = "mm-yy", values_to = "ttl") %>%
  mutate (Trade = substr(`mm-yy`,nchar(`mm-yy`),nchar(`mm-yy`))) %>%
  mutate (Date = substr(`mm-yy`,1,8)) %>%
  mutate(Trade= gsub("x","Export",Trade)) %>%
  mutate(Trade= gsub("y","Import",Trade)) %>%
  mutate(Date=gsub(" ","-",Date)) 

Chart2$Date<-ym(Chart2$Date)
Chart2 <- Chart2 %>%
  filter(`Data Series` %in% Countries) %>%
  mutate(`Data Series` = gsub("Federal Republic Of ()","",`Data Series`)) %>%
  mutate(`Data Series` = gsub("Socialist Republic Of ()","",`Data Series`))
Chart2 <- data.frame(Chart2,stringsAsFactors = TRUE)
```

```{r echo=FALSE}

base <- ggplot(Chart2, aes(Date, ttl, color=Trade)) + 
  geom_line() + 
  scale_color_manual (values=c("#00AFBB","#D16103")) +
  xlab(NULL) + 
  ylab("Total Value in SGD (million)") +
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))

p<-base + facet_wrap(Data.Series~.)

ggplotly(p)
```
