---
title: "Patterns of HDB Resale Prices by Residential Area in 2022"
subtitle: "Take Home Exercise 3"
author: "Wang Kunrui"
institute: "School of Computing and Information Systems,<br/>Singapore Management University"
date: "2020-2-15"
output:
  xaringan::moon_reader:
    css: [xaringan-themer.css, "css/text.css"]
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      beforeInit: "js/macros.js"
---

## Overview

HDB Prices are of interest to many Singaporeans; this document aims to provide insights on the difference in per square metre (PSM) rates for HDB sold in 2022.

# Data Wrangling

Using tidyverse family of packages, the following data wrangling was conducted: 1. Retrieved identifier for Year of Transaction, as the project is only interested in 2022 transactions 2. Filter down the dataset to 3-, 4- and 5- room flat types 3. Create the various "PSM" to calculate the per square metre rate at which the property was transacted

```{r, include=FALSE}
#| warning: false

library(pacman)
pacman::p_load(ggstatsplot, ggplot2, gapminder, ggside, knitr, tidyverse, broom, ggdist, gganimate, plotly, crosstalk, DT, hrbrthemes, viridis)
resale <- read_csv("resale-flat-prices-2017.csv")
resale <- transform(resale, yr= substr(month,1,4))
resale <- resale %>% filter(
  flat_type == "3 ROOM" | flat_type == "4 ROOM" | flat_type == "5 ROOM"
)
resale2022 <- resale %>% filter(
  yr == "2022"
  )
resale2022 <- transform(resale2022,
                        yrlease=(strtoi(substr(remaining_lease,1,2))))
resale2022 <- transform(resale2022,psm=resale_price/floor_area_sqm)
```

# A quick overview

In reaction to the rising resale prices post-covid, the government introduced property cooling measures on 30 September 2022.(See <a href="https://www.straitstimes.com/business/property/stricter-borrowing-criteria-for-property-buyers-15-month-wait-for-private-property-owners-moving-to-hdb-flats">this link</a> for details) The ruling imposed a 15 month waiting period for buyers who have sold their private home(s), before they are able to purchase a HDB flat. This is predominately expected to cool HDB resale demand.

A time-series plot on number of HDB resale transaction will provide insights on the cooling measure's impact across all 3 HDB flat types: 3-room, 4-room and 5-room.

```{r}

monthly <- resale2022 %>% count(month,flat_type)
p <- ggplot(monthly, aes(x=month, y=n, group=flat_type, color=flat_type, text=n)) +
  geom_line() + geom_point()+
  scale_color_viridis(discrete=TRUE) +
  theme(axis.text.x = element_text(angle=90))+
  ylab("Number of Transactions") 
ggplotly(p)

```

An immediate reaction is seen in the months of September and October; probably as buyers rushed to close deals ahead of implementation of rules. What is interesting - is the aftermath; In November and December, it appears that demand for 3- and 4- room HDB flat types are recovering back to Aug'22 levels.

More data points and future analysis will enrich this analysis; and help the public to understand the impact of these cooling measures.

# Deeper Analysis

Next, we seek to understand the difference purchase prices across HDB towns. Using ggplotly, an interactive boxplot was created; Users can hover across the box plot to observe the calculated transaction spread, including its various quartiles

```{r echo=FALSE}
#| warning: false
  
p <- ggplot(data=resale2022, aes(x=town, y=psm, fill=flat_type)) +
     geom_boxplot(width=0.1, alpha=0.2) +
     scale_fill_viridis(discrete = TRUE) +
     theme(axis.text.x = element_text(angle=90,size=8),
       plot.title = element_text(size=11))+
     ggtitle("Property price by sqm") +
   xlab ("TOWN")+ylab("SGD PER SQM")
 
 
ggplotly(p)
```

# Takeaways

Key observations are:

1.  Properties closer to Central region (e.g. Central Area, Bukit Merah, Queenstown, Kallang Whampoa, Toa Payoh) are sold at higher prices per sqm.

2.  In these same areas, we can observe a wider spread of prices per sqm, as demands for this area provide sellers more room for negotiation. The variance intra-town may be due to qualitative factors, including view from the home, complaints from neighbours, proximity to schools, etc.

3.  Interesting, we observe that some 5-room flats in area like Punggol and Sembawang are sold at lower prices per sqm as compared to 3- or 4- room flats in the same town. This may affected by factors such as remaining lease.
